### 2019-08-17 graphs/preamble.tex
Added Section 4

\usepackage{pgfplots}
\pgfplotsset{compat=1.16}
%\usepgfplotslibrary{external} 
%\tikzexternalize[prefix=pgfplots-cache/]

% Available ACM colours: ACMBlue, ACMYellow, ACMOrange, ACMRed, ACMLightBlue, ACMGreen, ACMPurple and ACMDarkBlue
\colorlet{GraphNormal}{ACMOrange} % Normal colour when I don't care about having multiple colours

\colorlet{GraphBase}{gray} % The colours for no type annotation
\colorlet{GraphA}{ACMRed}
\colorlet{GraphB}{ACMBlue}
\colorlet{GraphAB}{ACMPurple}

\pgfplotsset{%
% WARNING: do not put empty lines here, and make sure each = ... is terminated with a comma!
	every axis/.append style = {% options for all axes/graphs
		cycle list={}, tick align=center, xtick pos = bottom, ymin=0, axis on top,%
		axis y line*=left, try min ticks=5,%
		every tick label/.append style={font=\tiny},%
		every axis label/.append style={font=\tiny},%
	},%
	%
	Full/.style = { % For the page of graphs for each benchmark
		enlarge x limits={abs=1.5pt}, ymax={#1}, height=3.2cm, width=5.9cm,%
		every axis plot/.style={very thin, mark size=0.75pt, draw=GraphNormal, fill=GraphNormal!40}},%
	Right Axis/.style = { % For right axes
		axis y line*=right, axis x line=none, xmin=0, xmax=100},% 
	%
	Scatter/.style = { % For the scatter plots
		xticklabel={$\pgfmathprintnumber{\tick}$\%},%
		every axis plot/.append style={only marks}, mark=o},%
	%
	% Variants of the above that contain axis labels
	Scatter Left Label/.style   = {ylabel={Average Time (ms)}},%
	Scatter Bottom Label/.style = {xlabel={Proportion of Type Annotations}},%
	Scatter Right Label/.style  = {ylabel={Relative to 0\% Typed}},%
	%
	%
	Column/.style = {% For the column graphs
		enlarge x limits=false, xtick distance={#1}, const plot,%
		every axis plot/.append style={no markers}},%
	%
	% Variants of the above that contain axis labels
	Column Left Label/.style   = {ylabel={Average Time (ms)}},%
	Column Bottom Label/.style = {xlabel={Type Annotation Index}},%
	Column Right Label/.style  = {ylabel={Relative to Untyped}},%
	%
	Pattern/.style = { % For the page of graphs for each benchmark
		enlarge x limits={abs=1pt}, ymax={#1}, height=3.2cm, width=4.4cm, % 4.4
		every axis plot/.style={very thin, mark size=0.5pt, draw=GraphBase, fill=GraphBase!40}},%
	%
	Pattern A/.style = {draw=GraphA, fill=GraphA!40, mark=*}, %
	Pattern B/.style = {draw=GraphB, fill=GraphB!40, mark=*}, %
	Pattern AB/.style = {draw=GraphAB, fill=GraphAB!40, mark=*}, %
}

\newcommand\DOUBLEAXIS[3]{%
	\begin{axis}[#1]#3\end{axis}%
	\begin{axis}[#2]\end{axis}}

\newcommand\GRAPHS[1]{\null\hfill\begin{tikzpicture}[ampersand replacement=\&]\matrix{#1\\};\end{tikzpicture}\null\hfill}